### 2019-04-21 tikzpictures/flatten_matrix.tex
update some images

\begin{tikzpicture}
    \matrix [matrix of nodes,row sep=-1ex, column sep=-1ex, anchor=west] (toad_mat) at (0.0, 0.0)
    {
        5 & 7 & 10 & 10 & 11 & 4 \\
        6 & 8 & 8 & 10 & 10 & 5 \\
        5 & 6 & 8 & 8 & 7 & 4 \\
        6 & 5 & 6 & 5 & 7 & 5 \\
        6 & 6 & 4 & 5 & 4 & 4 \\
        5 & 5 & 4 & 2 & 3 & 6  \\  };

    \matrix [matrix of nodes,row sep=-1ex, column sep=-1ex, anchor=west] (toad_r1) at ([shift={(0.4, 0.00)}]toad_mat.east)
    {
        5 & 7 & 10 & 10 & 11 & 4 \\};

    \matrix [matrix of nodes,row sep=-1ex, column sep=-1ex, anchor=west] (toad_r2) at ([shift={(0.4, 0.00)}]toad_r1.east)
    {
        6 & 8 & 8 & 10 & 10 & 5 \\};

    \node (ellipsis) at  ([shift={(0.4, 0.00)}]toad_r2.east) {......};

    %\draw [->] ([shift={(0.05, 0.00)}]toad_mat.east ) -- ([shift={(-0.1, 0.00)}]toad_r1.west);
    %\draw [->] ([shift={(0.05, -0.60)}]toad_mat.east ) -- ([shift={(-0.1, 0.00)}]toad_r2.west);
    \draw [->] ([shift={(0.05, -0.40)}]toad_mat.north east) to [out=0, in=180] ([shift={(-0.1, 0.00)}]toad_r1.west);
    \draw [->] ([shift={(0.05, -0.75)}]toad_mat.north east) to [out=0, in=160] ([shift={(-0.1, 0.00)}]toad_r2.west);

\end{tikzpicture}