### 2019-05-09 tikzpictures/gaussian_kernel.tex
complete basic smoothing part

\begin{tikzpicture}
    \matrix [matrix of nodes,row sep=-1ex, column sep=-1ex, anchor=west] (kernel_ori) at (0.0, 0.0)
    {
        0.0030 & 0.0133 & 0.0219 & 0.0133 & 0.0030 \\
        0.0133 & 0.0596 & 0.0983 & 0.0596 & 0.0133 \\
        0.0219 & 0.0983 & 0.1621 & 0.0983 & 0.0219 \\
        0.0133 & 0.0596 & 0.0983 & 0.0596 & 0.0133 \\
        0.0030 & 0.0133 & 0.0219 & 0.0133 & 0.0030 \\    };

        \matrix [matrix of nodes,row sep=-1ex, column sep=-1ex, anchor=west] (kernel_tra) at ([shift={(0.4, 0.00)}]kernel_ori.east)
    {
        0.0030 & 0.0133 & 0.0219 & 0.0133 & 0.0030 \\
        0.0133 & 0.0596 & 0.0983 & 0.0596 & 0.0133 \\
        0.0219 & 0.0983 & \bf{0.0000} & 0.0983 & 0.0219 \\
        0.0133 & 0.0596 & 0.0983 & 0.0596 & 0.0133 \\
        0.0030 & 0.0133 & 0.0219 & 0.0133 & 0.0030 \\    };

        \matrix [matrix of nodes,row sep=-1ex, column sep=-1ex, anchor=west] (kernel_tra2) at ([shift={(0.4, 0.00)}]kernel_tra.east)
    {
        0.0035 & 0.0159 & 0.0262 & 0.0159 & 0.0035 \\
        0.0159 & 0.0712 & 0.1173 & 0.0712 & 0.0159 \\
        0.0262 & 0.1173 & 0.0000 & 0.1173 & 0.0262 \\
        0.0159 & 0.0712 & 0.1173 & 0.0712 & 0.0159 \\
        0.0035 & 0.0159 & 0.0262 & 0.0159 & 0.0035 \\
    };

    \draw [->] ([shift={(0.05, -0.00)}]kernel_ori.east) to [out=0, in=180] ([shift={(-0.1, 0.00)}]kernel_tra.west);
    \draw [->] ([shift={(0.05, -0.00)}]kernel_tra.east) to [out=0, in=180] ([shift={(-0.1, 0.00)}]kernel_tra2.west);

    \node at ([shift={(0.0, -0.2)}]kernel_ori.south) {Origin gaussian kernel matrix};
    \node at ([shift={(0.0, -0.2)}]kernel_tra.south) {Leave one};
    \node at ([shift={(0.0, -0.2)}]kernel_tra2.south) {Re-normalization};

\end{tikzpicture}